### 2022-06-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@b64f51898236a38c2da87a7aeaadb800649da1ab 🚀

\hypertarget{the-coevolutionary-mosaic-generates-emergence-regimes}{%
\subsection{The coevolutionary mosaic generates emergence
regimes}\label{the-coevolutionary-mosaic-generates-emergence-regimes}}

@@

Based on the geographic mosaic theory of coevolution, we developed a
trivariate map of three facets of coevolutionary pressure (see Methods):
(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
histories should expose viruses to more variation in host immune traits;
(2) \emph{host community uniqueness}: coevolving with more unique host
communities should create more unique branches of viral evolution; and
(3) propensity for \emph{viral sharing}: frequent cross-species
transmission may act as a buffer on selective pressure, while lower
rates of exchange may enable more simultaneous trajectories of viral
specialization to coexist within a given community. We combine global
maps of all three to generate a map of coevolutionary regimes, where
close colors represent similar risks, and paler pixels represent overall
higher risk (fig.~\ref{fig:trivariate}). We find that these regions do
not neatly overlap with those defined in fig.~\ref{fig:richness} or
fig.~\ref{fig:biogeo}, reinforcing the notion that local-scale
coevolutionary mosaics can form within cophylogenetic regions.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{Trivariate additive mapping of the components of risk in the
red/green/blue, where high virus sharing is encoded in the blue channel,
host phylogenetic diversity in the green channel, and compositional
uniqueness in the red channel. The main driver of emergence risk
(possibility of spatially overlapped but coevolutionarily indepent
host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
pixels around yellow. Pixels in the yellow-green space (Africa)
correspond to areas where, despite the potential for viral
diversification, coevolutionary dynamics are likely to currently be
dominated by the effect of host phylogenetic
diversity.}\label{fig:trivariate}
}
\end{figure}

Emergence risk is maximized under low viral sharing (host-virus pairs
coevolve independantly), high phylogenetic diversity (viruses are
exposes to different host clades), and high host uniqueness (viruses are
experiencing novel, heterogeneous host traits combinations). Under these
conditions, very different betacoronaviruses could co-exist at the same
place, yet evolve in independent ways. As betacoronaviruses often evolve
(including host shifts) through recombination, the co-occurrence of
sufficiently distinct viruses is a sufficient major driver of emergence,
and the regions that meet these conditions therefore represent the
higher risk. In fig.~\ref{fig:trivariate}, this corresponds to yellow
areas (dynamics dominated by low viral sharing, with equal contributions
of selection mosaics and trait remixing; South-Eastern Asia, and the
Indian sub-continent), green-yellow areas (dynamics with low viral
sharing but dominated by the selection mosaic effect of host diversity;
Africa below the Sahara desert), and would correspond to red-yellow
areas (dynamics with low viral sharing but dominated by trait remixing
in host communities; Middle-East). Indeed, these regions are broadly
reflected in the risk map fig.~\ref{fig:risk}.

Under this framework, other regions (where high viral sharing dominates
the dynamics; see \emph{e.g.} Latin America, Eurasia above a northing of
30) represent a lower risk of emergence. Nevertheless, areas of high
host uniqueness coupled with high viral sharing(red-to-pink in
fig.~\ref{fig:trivariate}) could provide future hotspots of
\emph{Betacoronavirus} emregence risk through the suddent divergence of
currently diffuse coevolutionary dynamics; this is a likely scenario
knowing that can be facilitated by codivergence followed by
recombination. Madagascar, where most bat species are endemic following
evolutionary divergence from sister species in both African and Asian
continents,\textsuperscript{24} is one such potential future hotspot.
Indeed, recent surveillance\textsuperscript{25} has identified a novel
\emph{Betacoronavirus} (in the subgenus \emph{Nobecovirus}) in
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus},
\emph{Rousettus madagascariensis}), emphasizing strong proof of
principle in model predictions.

Our risk decomposition does not account for viral diversity or
distinctiveness. Viral data acquisition is rarely disconnected from the
acquisition of host data. There are more sources of information on hosts
than on viruses, allowing to develop a more robust host-centric
perspective on risk. Any emergence risk estimate would benefit from
viral traits related to \emph{e.g.} ability to switch hosts or
pathogenic potential. This is particularly true under recent findings
that the diversification of bat coronaviruses is driven largely by host
shifts (inter-genus or higher levels of cross-species transmission) and,
to a lesser degree, cospeciation and sharing, representing intra-genus
cross-species transmission.\textsuperscript{2} This diversification is
not an actual risk factor for emergence itself, but acts downstream of
an emergence event by increasing the random chance of the emergence of a
virus with the raw genomic components required to eventually infect
humans.

>>>>

\hypertarget{host-landscapes-generate-a-coevolutionary-mosaic}{%
\subsection{Host landscapes generate a coevolutionary
mosaic}\label{host-landscapes-generate-a-coevolutionary-mosaic}}

@@

Based on the geographic mosaic theory of coevolution, we developed a
trivariate map of three facets of coevolutionary pressure (see Methods):
(1) \emph{host phylogenetic diversity}: a high diversity of evolutionary
histories should expose viruses to more variation in host immune traits;
(2) \emph{host community uniqueness}: exposure to greater host trait
heterogeneity can drive viral diversification, and coevolving with more
unique host communities should create more unique branches of viral
evolution; and (3) propensity for \emph{viral sharing}: frequent
cross-species transmission may act as a buffer on selective pressure,
while lower rates of exchange may enable more simultaneous trajectories
of viral specialization to coexist within a given community. We combine
global maps of all three to generate a map of coevolutionary regimes,
where close colors represent similar risks, and paler pixels represent
overall higher risk (fig.~\ref{fig:trivariate}). We find that these
regions do not neatly overlap with those defined in
fig.~\ref{fig:biogeo} or fig.~\ref{fig:richness}, reinforcing the notion
that local-scale coevolutionary mosaics can form within cophylogenetic
regions.

\begin{figure}
\hypertarget{fig:trivariate}{%
\centering
\includegraphics{figures/risk_trivariate.png}
\caption{Trivariate additive mapping of the components of risk in the
red/green/blue, where high virus sharing is encoded in the blue channel,
host phylogenetic diversity in the green channel, and compositional
uniqueness in the red channel. The main driver of emergence risk
(possibility of spatially overlapped but coevolutionarily indepent
host-viral dynamics) corresponds to low viral sharing, \emph{i.e.}
pixels around yellow.}\label{fig:trivariate}
}
\end{figure}

The greatest evolutionary potential for zoonotic emergence exists where
pathogen pools have a high genetic diversity and high propensity for
cross-species transmission. In our framework, emergence risk is
therefore maximized under higher phylogenetic diversity (viruses are
exposes to different host clades), higher host uniqueness (viruses are
experiencing novel, heterogeneous host traits combinations), and low to
medium viral sharing (host-virus pairs can coevolve independently, but
divergent viruses may still have opportunities for recombination). In
fig.~\ref{fig:trivariate}, this corresponds to yellow areas (dynamics
dominated by low viral sharing, with equal contributions of selection
mosaics and trait remixing; southeast Asia, and the Indian
sub-continent), green-yellow areas (dynamics with low viral sharing but
dominated by the selection mosaic effect of host diversity; Africa below
the Sahara desert), and red-yellow areas (dynamics with low viral
sharing but dominated by trait remixing in host communities; the Middle
East). Translating this axis of variation back into a univariate risk
map (fig.~\ref{fig:risk}) highlights that this evolutionary landscape
has a striking correspondence to regions where zoonotic
betacoronaviruses have previously emerged.

Under this framework, other regions where high viral sharing dominates
coevolutionary dynamics---such as Latin America, or Eurasia above a
northing of 30---would pose less of a risk of zoonotic emergence.
Nevertheless, areas of high host uniqueness coupled with high viral
sharing (red-to-pink in fig.~\ref{fig:trivariate}) could create hotspots
facilitated by viral codivergence. Our framework identifies Madagascar,
where most bat species are endemic following evolutionary divergence
from sister species in both African and Asian
continents,\textsuperscript{24} as one such hotspot; interestingly, a
recent study\textsuperscript{25} reported a novel and highly divergent
lineage of nobecoviruses from Madagascar-endemic pteropid bat species
(\emph{Pteropus rufus} and \emph{Rousettus madagascariensis}), again
supporting the predictive power of the coevolutionary framework.
